---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Preface
In this report, I explored and visualized some patterns of steps taken per day.

## Load library
```{r load library, echo = TRUE}
library(dplyr)
library(ggplot2)
```


## Loading and preprocessing the data
```{r load data, echo = TRUE}
act <- read.csv("./activity.csv")
Act <- tbl_df(act)
```
Explore the dataset
```{r explore dataset, echo =TRUE}
str(Act)
head(Act)
summary(Act)
```

## What is mean total number of steps taken per day?
### Creat tbl object for calculating the total steps per day
use sum function and group by date
```{r prepare, echo = TRUE}
table <- Act %>%
  group_by(date) %>%
  summarise(total_step=sum(steps))

# explore table
table
```

### Plot the histogram of the total steps taken each day
use base plot system  
```{r hist plot, echo = TRUE}
hist(table$total_step, breaks = 50, main = "Histogram of Total Steps Per Day",
     xlab = "total steps")
```

### Calculate the mean and median
```{r mean&median, echo = TRUE}
step_mean <- as.integer(mean(table$total_step, na.rm = TRUE))
step_median <- as.integer(median(table$total_step, na.rm = TRUE))
```
The mean of total steps taken per day is `r step_mean`, and the median is `r step_median`.

## What is the average daily activity pattern?  
### Prepare table for average steps across days of every 5-min interval
```{r prepare2, echo = TRUE}
table2 <- Act %>%
group_by(interval) %>%
summarise(aver_step=mean(steps, na.rm=TRUE))
```

### Plot {daily activity pattern.
```{r plot daily activity, echo = TRUE}
qplot(interval, aver_step, data = table2, geom = "line")
```

### Find the interval with max steps taken.
```{r find max, echo = TRUE}
max_interval <- which.max(table2$aver_step)
```
the interval with max steps taken is the `r max_interval`th(identifier: `r table2$interval[max_interval]`) interval.  

## Imputing missing values
### Count the NA values in the dataset
```{r count NA values, echo = TRUE}
count = sum(is.na(act$steps))
```
There is `r count` NA values in the dataset.  

### Find pattern of NA values
```{r find pattern of NA, echo = TRUE}
head(table(act$interval, is.na(act$steps)))
head(table(act$date, is.na(act$steps)))
```
So, the missing values are related to date.

### Use mean of 5-min interval to fill the NA values
```{r impute missing values, echo = TRUE}
# find the date whose record is missed
missingdate <- unique(subset(act,is.na(act$steps))$date)

# replace missing values in new dataset
act2<-act
act2$steps = replace(act$steps, which(act$date %in% as.character(missingdate)),
rep(table2$aver_step, times = length(missingdate)))

# verify the replacement
summary(act2)
```

### Make histogram, calculate the new mean and median
```{r make histogram & calculate mean and median, echo = TRUE}
Act2 <- tbl_df(act2)

# prepare for plot and calculation
table3 <- Act2 %>%
  group_by(date) %>%
  summarise(total_step=sum(steps))

# plot histogram
hist(table3$total_step, breaks = 50, main = "Histogram of Total Steps Per Day",
     xlab = "total steps")
step_mean2 <- as.integer(mean(table3$total_step))
step_median2 <- as.integer(median(table3$total_step))
```
The mean of total steps taken per day is `r step_mean2`, and the median is `r step_median2`.   

### Coclusion
Compared with those with missing values included, it seems mean and median changed slightly, and the frequency is more likely normal distributed.

## Are there differences in activity patterns between weekdays and weekends?
### Find weekends in dates give
```{r find weekends, echo = TRUE}
dates<-as.character(unique(act$date))
weekends<-dates[weekdays(as.Date(dates)) == "星期日"| weekdays(as.Date(dates)) == "星期六"]
```
### Create new variables to identify weekdays and weekends
```{r identify weekdays and weekends, echo = TRUE}
Act2 <- mutate(Act2, weekday = (date %in% weekends))
Act2$weekday = factor(Act2$weekday, labels = c("weekday", "weekend"))
```
### Plot comparison between weekdays and weekends
```{r plot comparison, echo = TRUE}
# prepare table
table4 <- Act2 %>%
group_by(interval, weekday) %>%
summarise(aver_step=mean(steps))

## make plot
g <- ggplot(data = table4)
g + geom_line(aes(x = interval, y = aver_step)) + facet_wrap(~weekday, nrow = 2)
```

### Conclusion
There are more steps taken in weekends than in weekdays.